Influence of anomalous perpendicular transport on linear tearing mode dynamics in tokamak plasmas

### 2021-12-11 Research/LinearTearingModeTheory/LinearTearingModeTheory.tex
Started writing linear and nonlinear papers

\title{Further drift-magnetohydrodynamic modeling of linear tearing mode dynamics in tokamak plasmas}

@@

\begin{abstract}
An improved 
\end{abstract}

@@

\section*{Acknowledgements}
This research was directly funded by the U.S.\ Department of Energy, Office of Science, Office of Fusion Energy Sciences,  under contract DE-FG02-04ER54742,

>>>>

\title{Influence of anomalous perpendicular transport on linear tearing mode dynamics in tokamak plasmas}

@@

\begin{abstract}
The analysis of a previous paper [A.~Cole, and R.~Fitzpatrick, Phys.\ Plasmas {\bf 13}, 032503 (2006)] that maps out all of the
two-fluid response regimes of a  linear tearing layer interacting with an externally generated resonant magnetic perturbation, in a large aspect-ratio tokamak plasma,  is generalized to
take into account realistic levels of perpendicular particle transport. A new response-regime map is obtained that differs substantially, in many
respects, from the old one. The improved analysis is first employed to find all of the two-fluid growth-rate regimes of a non-interacting low-mode-number tearing mode in a large aspect-ratio tokamak plasma. 
The analysis is then used to determine the scaling of the error-field penetration threshold with machine
parameters in large aspect-ratio, ohmically heated, tokamak plasmas . A number of penetration regimes are found in which the threshold scales as $n^{1/2}$, where $n$ is the line-averaged
electron number density. A similar scaling has been reported in some experiments. 
\end{abstract}

@@

\section{Introduction}
Tearing modes are slowly growing instabilities of ideally-stable tokamak plasmas that reconnect magnetic field-lines
at various resonant surfaces within the plasma, in the process forming magnetic island chains that degrade the plasma confinement.\cite{wes}
If tearing modes grow to sufficiently large amplitude then they can trigger major disruptions.\cite{wes1}  Tokamak
plasmas are observed to be particularly disruption-prone when tearing modes {\em lock}\/ (i.e., become stationary in the
laboratory frame) to externally generated, resonant magnetic perturbations.\cite{vries}

It is well known that single-fluid, resistive magnetohydrodynamics (MHD) offers a very poor description of
tearing mode dynamics in tokamak plasmas. 
For instance, the strong {\em diamagnetic}\/ flows present in such plasmas decouple the electron and ion flows to some extent,
necessitating a two-fluid treatment.\cite{ara} Moreover, resistive-MHD does not take  the important {\em ion sound radius lengthscale}, below which electron and ion dynamics are further decoupled, 
into account.\cite{drake,wal} Previously, Cole \& Fitzpatrick\,\cite{cole} used the four-field model
of Fitzpatrick \& Waelbroeck\,\cite{fw} (which is based on the original four-field model of Hazeltine, et al.\cite{haz}) 
to determine the two-fluid response of a linear tearing layer to an externally generated, 
resonant magnetic perturbation  in a large aspect-ratio tokamak plasma. However, the treatment of Cole \& Fitzpatrick fails to take into account the
 anomalously large perpendicular  particle diffusivity  present in tokamak plasmas. The
aim of this paper is to correct this deficiency.

This paper is organized as follows. After some preliminary analysis in Sect.~\ref{sect1}, our fundamental four-field resonant
plasma response model is introduced in Sect.~\ref{sfour}. This model is a slightly improved version of the model derived in Ref.~\onlinecite{fw} that takes the thermal force into account. The model is converted into a set of linear layer equations in
Sect.~\ref{layer}. These layer equations are then solved in Sect.~\ref{linear} in order to map out all of the linear response regimes
of a resonant layer to an externally generated resonant magnetic perturbation in a large aspect-ratio tokamak plasma. In Sect.~\ref{stab}, we reuse the analysis of Sect.~\ref{linear} to map out all
of the growth-rate regimes for a low-mode-number (but not $m=1$) tearing mode in a large aspect-ratio tokamak plasma. Finally,
in Sect.~\ref{error}, we employ the results of Sect.~\ref{linear} to determine the scaling of the error-field penetration threshold with machine parameters
 in large aspect-ratio, ohmically heated, tokamak plasmas.

\section{Preliminary Analysis}\label{sect1}
\subsection{Plasma Equilibrium}
Consider a large aspect-ratio tokamak plasma equilibrium whose magnetic flux-surfaces map out
(almost) concentric circles in the poloidal plane. Such an equilibrium can be approximated as a
periodic cylinder.\cite{rf1993} Let $r$, $\theta$, $z$ be right-handed cylindrical coordinates. 
The magnetic axis corresponds to $r=0$, and the plasma boundary to $r=a$. The system is assumed to be periodic in the $z$
direction with periodicity length $2\pi\,R_0$, where $R_0$ is the simulated major radius of the plasma. The
safety-factor profile takes the form $q(r)=r\,B_z/[R_0\,B_\theta(r)]$, where $B_z$ is the constant
`toroidal' magnetic field-strength, and $B_\theta(r)$ is the poloidal magnetic field-strength. The standard
large aspect-ratio orderings, $r/R_0\ll 1$ and $B_\theta/B_z\ll 1$, are adopted.

\subsection{Asymptotic Matching}\label{cyl}
Consider a tearing mode perturbation that has $m$ periods in the poloidal direction, and
$n$ periods in the toroidal direction. 
As is well known, the response of the plasma to the tearing mode is
governed by marginally stable, ideal-MHD everywhere in the plasma, apart from a (radially) narrow
layer centered on the so-called {\em resonant surface}, minor radius $r_s$, at which $q(r_s)=m/n$.\cite{fkr}

The perturbed magnetic field associated with the tearing mode is written $\delta{\bf B} \simeq \nabla\delta\psi\times{\bf e}_z$,  
where
$\delta\psi(r,\theta,\varphi,t)= \delta\psi(r,t)\,\exp[\,{\rm i}\,(m\,\theta-n\,\varphi)]$, 
and $\varphi=z/R_0$ is a simulated toroidal angle.

In the so-called {\em outer region}\/ (i.e., everywhere in the plasma
apart from the resonant layer), the perturbed helical magnetic flux, $\delta\psi(r,t)$, satisfies the
{\em cylindrical tearing mode equation}:\cite{wes}
\begin{equation}\label{e3}
\frac{\partial^2\delta\psi}{\partial r^2} + \frac{1}{r}\,\frac{\partial\delta\psi}{\partial r}-\frac{m^2}{r^2}\,\delta\psi - \frac{J_z'\,\delta\psi}{r\,(1/q-n/m)}=  0,
\end{equation}
where 
$J_z(r)= R_0\,\mu_0\,j_z(r)/B_z$,
and $j_z(r)$ is the equilibrium `toroidal' current density. Here, $'\equiv d/dr$. In general, the solution of Eq.~(\ref{e3}) that satisfies physical
boundary conditions at the magnetic axis and the plasma boundary is such that $\delta\psi$ is continuous
across the resonant surface, whereas $\partial\delta\psi/\partial r$ is discontinuous. The discontinuity of
 $\partial\delta\psi/\partial r$ across the resonant surface is indicative of the presence of a helical current
 sheet at the surface.  The complex quantity ${\mit\Psi}_s(t)= \delta\psi(r_s,t)$ determines the amplitude
 and phase of the reconnected helical magnetic flux at the resonant surface, whereas the complex quantity\,\cite{rf1993}
 \begin{equation}
{\mit\Delta\Psi}_s = \left[r\,\frac{\partial \delta\psi}{\partial r}\right]_{r_{s-}}^{r_{s+}}
\end{equation}
parameterizes the amplitude and phase of the helical current sheet flowing at the surface. The solution of
the cylindrical tearing mode equation in the outer region, in the presence of an externally generated, resonant magnetic perturbation 
of the same helicity as the tearing mode, leads to the relation\,\cite{rf1993,fkr}
\begin{equation}\label{e110}
{\mit\Delta\Psi}_s= {\mit\Delta}_s'\,{\mit\Psi}_s + (-{\mit\Delta}_s')\,{\mit\Psi}_v,
\end{equation}
where ${\mit\Delta}_s'$ is a real dimensionless quantity known as the {\em tearing stability index}. Moreover,
${\mit\Psi}_v$ is the so-called {\em vacuum flux}, and is defined as the reconnected magnetic flux
that would be driven at the resonant surface by the resonant magnetic perturbation were the
plasma intrinsically tearing stable (i.e., ${\mit\Delta}_s' <0$), and were there no current sheet at the resonant surface
(i.e., ${\mit\Delta\Psi}_s=0$).

The current sheet at the resonant surface can only be resolved by solving  a two-fluid, resistive-MHD, plasma response model
in the so-called {\em inner region}\/ (i.e., the region of the plasma in the immediate vicinity of the resonant surface), 
and asymptotically matching the solution so obtained to the ideal-MHD solution in the outer region. The particular 
plasma response model used in this paper is
described in the following section.

\section{Four-Field Resonant Plasma Response Model}\label{sfour}
\subsection{Useful Definitions}
The plasma is assumed to consist of two species. First, electrons of mass $m_e$, electrical charge $-e$, 
number density $n$, and temperature $T_e$.  Second, ions of mass $m_i$, electrical charge $+e$,  
number density $n$, and temperature $T_i$. Let $p=n\,(T_e+T_i)$ be the total plasma pressure.

It is helpful to define $n_0 = n(r_s)$, $p_0= p(r_s)$,
\begin{align}
\eta_e &=\left.\frac{d\ln T_e}{d\ln n}\right|_{r=r_s},\label{e211}\\[0.5ex]
\eta_i &= \left.\frac{d\ln T_i}{d\ln n}\right|_{r=r_s},\\[0.5ex]
\tau &= \left(\frac{T_e}{T_i}\right)_{r=r_s}\left(\frac{1+\eta_e}{1+\eta_i}\right),\label{e213}
\end{align}
where $n(r)$, $p(r)$, $T_e(r)$, and $T_i(r)$ refer to
density, pressure, and temperature profiles that are unperturbed by the tearing mode.

For the sake of simplicity, the perturbed electron and ion temperature profiles are assumed to be functions of
the perturbed electron number density profile in the immediate vicinity of the resonant surface. In other words, $T_e=T_e(n)$ and $T_i=T_i(n)$. This
implies that $p=p(n)$. 
The {\em MHD velocity}, which is the velocity of a
fictional MHD fluid, is defined ${\bf V}={\bf V}_E + V_{\parallel\,i}\,{\bf b}$, where ${\bf V}_E$ is the
${\bf E}\times{\bf B}$ drift velocity, $V_{\parallel\,i}$ is the parallel component of the ion fluid
velocity, ${\bf b}= {\bf B}/|{\bf B}|$, and ${\bf B}$ is the magnetic field-strength.

\subsection{Fundamental Fields}
The four fundamental fields in our four-field model---namely, $\psi$, $\phi$, $N$, and $V$---have the following
definitions:
\begin{align}\label{e10}
\nabla\psi &= \frac{{\bf n}\times{\bf B}} {r_s\,B_z},\\[0.5ex]
\nabla\phi &= \frac{{\bf n}\times {\bf V}}{r_s\,V_A},\\[0.5ex]
N &=-\hat{d}_i\left(\frac{p-p_0}{B_z^{\,2}/\mu_0}\right),\\[0.5ex]
V &= \hat{d}_i\left(\frac{{\bf n}\cdot {\bf V}}{V_A}\right).\label{e13}
\end{align}
Here,   ${\bf n} = (0,\,\epsilon/q_s,\,1)$, $\epsilon = r/R_0$, $q_s=m/n$, 
$V_A =B_z/\sqrt{\mu_0\,n_0\,m_i}$, 
$d_i = \sqrt{m_i/(n_0\,e^{\,2}\,\mu_0)}$,
and $\hat{d}_i=d_i/r_s$. 
 Our
model also employs the auxiliary field
\begin{align}\label{e16}
J=-\frac{2\,\epsilon_s}{q_s}+\hat{\nabla}^2\psi,
\end{align}
where 
$\epsilon_s=r_s/R_0$, and $\hat{\nabla} = r_s\,\nabla$.

\subsection{Fundamental Equations}
The four-field model takes the form:\cite{fw}
\begin{align}
\frac{\partial\psi}{\partial\hat{t}}&= [\phi,\psi] -\left(\frac{\tau}{1+\tau}\right)(1+\lambda_e)\,[N,\psi]
+\hat{\eta}_\parallel\,J + \hat{E}_\parallel,\\[0.5ex]
\frac{\partial \hat{\nabla}^2\phi}{\partial \hat{t}}&= [\phi,\hat{\nabla}^2\phi] - \frac{1}{2\,(1+\tau)}\left(\hat{\nabla}^2[\phi,N] + [\hat{\nabla}^2\phi,N] + [\hat{\nabla}^2 N,\phi]\right) + [J,\psi] \nonumber\\[0.5ex]&\phantom{=}+\hat{\chi}_\varphi  \,\hat{\nabla}^4\!\left(\phi + \frac{N}{1+\tau}\right), \\[0.5ex]
\frac{\partial N}{\partial \hat{t}}&= [\phi,N] +c_\beta^{\,2}\,[V,\hat{\psi}] +\hat{d}_\beta^{\,2}\,[J,\psi]
+ \hat{D}_\perp\,\hat{\nabla}_\perp^{\,2}N,\\[0.5ex]
\frac{\partial V}{\partial\hat{t}}&= [\phi,V] +[N,\psi] + \hat{\chi}_\varphi\,\hat{\nabla}^2 V.\label{e21}
\end{align}
Here, $[A,B]\equiv \hat{\nabla} A\times \hat{\nabla} B\cdot {\bf n}$, $\hat{t} = t/(r_s/V_A)$, $\hat{\eta}_{\parallel,\perp} = \eta_{\parallel,\perp}/(\mu_0\,r_s\,V_A)$, $\hat{E}_\parallel = E_\parallel/(B_z\,V_A)$, 
$\hat{\chi}_\varphi= \chi_\varphi/(r_s\,V_A)$, where $\eta_{\parallel,\perp}$ is the parallel/perpendicular plasma electrical
resistivity at the resonant surface, $E_\parallel$ the parallel inductive electric field that maintains the equilibrium toroidal
plasma current in the vicinity of the resonant surface, and $\chi_\varphi$  the anomalous perpendicular ion momentum
diffusivity at the resonant surface. 
Moreover, $d_\beta=c_\beta\,d_i$, and $\hat{d}_\beta=d_\beta/r_s$, where $c_\beta = \sqrt{\beta/(1+\beta)}$, and
$\beta=(5/3)\,\mu_0\,p_0/B_z^{\,2}$. Here, $d_\beta$ is usually referred to as the {\em ion sound radius}. Furthermore,
$\lambda_e=0.71\,\eta_e/(1+\eta_e)$. Finally,\cite{fw}
\begin{equation}
\hat{D}_\perp \equiv \frac{D_{\perp}}{r_s\,V_A}= c_\beta^{\,2}\,\hat{\eta}_\perp +  \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\perp\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\perp\,i}\right),
\end{equation}
where $\hat{\chi}_{\perp\,e,i}=\chi_{\perp\,e,i}/(r_s\,V_A)$, and $\chi_{\perp\,e,i}$ is the anomalous
electron/ion perpendicular energy diffusivity in the vicinity of the resonant surface. 
Note that Cole \& Fitzpatrick (2006) neglects the contribution of $\chi_{\perp\,e,i}$ to the perpendicular particle
diffusivity, $D_\perp$, leading to an anomalously small estimate for $D_\perp$. It is this deficiency that we aim to
correct in the present paper.

\subsection{Matching to Plasma Equilibrium}
The unperturbed plasma equilibrium is such that
${\bf B} = (0,\,B_\theta(r),\,B_z)$,  $p = p(r)$,
${\bf V} = (0,\,V_E(r),\,V_z(r))$,
where 
$V_E(r)\simeq E_r/B_z$
 is the (dominant $\theta$-component of the) ${\bf E}\times {\bf B}$ velocity. Now, the resonant layer is assumed to have a radial thickness that is
much smaller than $r_s$.   Hence, we only need to evaluate plasma equilibrium quantities in the immediate vicinity of the resonant
surface. Equations~(\ref{e10})--(\ref{e13}) suggest that 
\begin{align}
\psi&= \frac{\hat{x}^{\,2}}{2\,\hat{L}_s},\label{e23}\\[0.5ex]
\phi &= - \hat{V}_E\,\hat{x},\\[0.5ex]
N &= -\hat{V}_\ast\,\hat{x},\\[0.5ex]
V &= \hat{V}_\parallel,\label{e26}
\end{align}
where 
$\hat{x}=(r-r_s)/r_s$,
 $\hat{L}_s=L_s/r_s$,  $L_s=R_0\,q_s/s_s$, 
  $\hat{V}_E= V_E(r_s)/V_A$,
$\hat{V}_\ast= V_\ast(r_s)/V_A$,
$V_\ast(r) = (dp/dr)/(e\,n_0\,B_z)$ 
is the (dominant $\theta$-component of the) diamagnetic velocity,
  and 
 $\hat{V}_\parallel=\hat{d}_i\, V_z(r_s)/V_A$. Here, $s_s=s(r_s)$ and $s(r)=d\ln q/d\ln r$. We also have
 \begin{align}\label{e28}
 J &= -\left(\frac{2}{s_s}-1\right)\frac{1}{\hat{L}_s},
\end{align} 
and $ \hat{E}_\parallel =(2/s_s-1) (\hat{\eta}_\parallel/\hat{L}_s)$.

\section{Linear Resonant Plasma Response Model}\label{layer}
\subsection{Introduction}
The aim of this section is to obtain a set of linear layer equations from the four-field model
introduced in Sect.~\ref{sfour}.

\subsection{Derivation of Linear Layer Equations}
In accordance with Eqs.~(\ref{e23})--(\ref{e28}), let us write
\begin{align}\label{e31}
\hat{\psi}(\hat{x},\zeta,\hat{t}) &= \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ \tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
\phi(\hat{x},\zeta,\hat{t}) &=-\hat{V}_E\,\hat{x}+ \tilde{\phi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
N(\hat{x},\zeta,\hat{t}) &= -\hat{V}_\ast\,\hat{x}+\left(\frac{1+\tau}{\tau}\right)\!\tilde{N}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
V(\hat{x},\zeta,\hat{t}) &= \hat{V}_\parallel + \left(\frac{1+\tau}{\tau}\right)\!\tilde{V}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
J(\hat{x},\zeta,\hat{t}) &=-\left(\frac{2}{s_s}-1\right)\!\frac{1}{\hat{L}_s}+ \hat{\nabla}^2\tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\label{e36}
\end{align}
where $\zeta=m\,\theta-n\,\varphi$, $\hat{\omega}=r_s\,\omega/V_A$, and $\omega$ is the frequency of the tearing mode in the laboratory frame. 
Substituting Eqs.~(\ref{e31})--(\ref{e36}) into Eqs.~(\ref{e16})--(\ref{e21}), 
 and only retaining terms that
are first order in perturbed quantities, we obtain the following set of linear equations:
\begin{align}\label{e37}
-{\rm i}\,[\omega-\omega_E -(1+\lambda_e)\,\omega_{\ast\,e}]\,\tau_H\,\tilde{\psi} &= -{\rm i} \,\hat{x}\,[\tilde{\phi}-(1+\lambda_e)\,\tilde{N}] + S^{-1}\,\hat{\nabla}^2\tilde{\psi},\\[0.5ex]
-{\rm i}\,(\omega-\omega_E-\omega_{\ast\,i})\,\tau_H\,\hat{\nabla}^2\tilde{\phi} &= -{\rm i}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi}+S^{-1}\,P_\varphi\,\hat{\nabla}^4\!\left(\tilde{\phi} + \frac{\tilde{N}}{\tau}\right),\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{N} &= -{\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\phi}-{\rm i}\,c_\beta^{\,2}\,\hat{x}\,\tilde{V} - {\rm i}\!\left(\frac{\tau}{1+\tau}\right)\hat{d}_\beta^{\,2}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi}\nonumber\\[0.5ex]
&\phantom{=} + S^{-1}\,P_\perp\,\hat{\nabla}_\perp^2 \tilde{N},\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{V} &= {\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\psi} - {\rm i}\,\hat{x}\,\tilde{N}
+ S^{-1}\,P_\varphi\,\hat{\nabla}^2\tilde{V}.\label{e40}
\end{align}
Here, 
$\tau_H = L_s/(m\,V_A)$ 
is the  hydromagnetic time, 
$\omega_E =(m/r_s)\,V_E(r_s)$
  the 
 ${\bf E}\times {\bf B}$ frequency, 
$\omega_{\ast\,e} = -[\tau/(1+\tau)](m/r_s)V_\ast(r_s)$
the 
   electron diamagnetic
frequency,
$\omega_{\ast\,i} =[1/(1+\tau)](m/r_s)V_\ast(r_s)$
 the  ion diamagnetic
frequency, 
$S=\tau_R/\tau_H$
 the  Lundquist number, 
$\tau_R = \mu_0\,r_s^{\,2}/\eta_\parallel$
 the
 resistive diffusion time, 
$\tau_\varphi
= r_s^{\,2}/\chi_\varphi$
the  toroidal momentum confinement time, and 
$\tau_\perp = r_s^{\,2}/D_\perp$.
  Furthermore, $P_\varphi = \tau_R/\tau_\varphi$ and $P_\perp = \tau_R/\tau_\perp$ are magnetic Prandtl numbers.

 Let us define the stretched radial variable $X = S^{1/3}\,\hat{x}$.
Assuming that $X\sim{\cal O}(1)$ in the layer (i.e., assuming that the layer thickness is roughly of order $S^{-1/3}\,r_s$),
and making use of the fact that $S\gg 1$ in conventional tokamak plasmas,  Eqs.~(\ref{e37})--(\ref{e40}) reduce to the following
set of linear layer equations:\,\cite{cole}
\begin{align}\label{e43}
-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_{e}]\,\tilde{\psi}&= - {\rm i}\,X\left[\tilde{\phi}-(1+\lambda_e)\,\tilde{N}\right] + \frac{d^{\,2}\tilde{\psi}}{d X^2},\\[0.5ex]
-{\rm i}\,(Q-Q_E -Q_{i})\,\frac{d^{\,2}\tilde{\phi}}{dX^2}&= - {\rm i}\,X\,\frac{d^{\,2}\tilde{\psi}}{dX^2}+ P_\varphi\,\frac{d^{\,4}}{dX^4}\!\left(\tilde{\phi} + \frac{\tilde{N}}{\tau}\right),\\[0.5ex]
-{\rm i}\,(Q-Q_E)\tilde{N} &= - {\rm i}\,Q_{e}\,\tilde{\phi} - {\rm i}\,c_\beta^{\,2}\,X\,\tilde{V}  - {\rm i} \,D^{2}\,X\,\frac{d^{\,2}\tilde{\psi}}{dX^{2}}
+ P_\perp\,\frac{d^{\,2} \tilde{N}}{dX^{2}},\\[0.5ex]
 -{\rm i}\,(Q-Q_E)\,\tilde{V} &= {\rm i}\,Q_{e}\,\tilde{\psi} - {\rm i}\,X\,\tilde{N} + P_\varphi\,\frac{d^{\,2}\tilde{V}}{dX^{2}}.\label{e46}
\end{align}
Here, $Q=S^{\,1/3}\,\omega\,\tau_H$, $Q_E = S^{\,1/3}\,\omega_E\,\tau_H$, $Q_{e,i} = S^{\,1/3}\,\omega_{\ast\,e,i}\,\tau_H$,
and $D = S^{\,1/3}\,[\tau/(1+\tau)]^{1/2}\,\hat{d}_\beta$.

\subsection{Asymptotic Matching}
The  linear layer equations, (\ref{e43})--(\ref{e46}), possess tearing parity solutions
characterized by the symmetry $\tilde{\psi}(-X)=\tilde\psi(X)$, $\tilde{\phi}(-X)=-\tilde{\phi}(X)$, 
$\tilde{N}(-X)= - \tilde{N}(X)$, $\tilde{V}(-X)=\tilde{V}(X)$. If we assume that
the asymptotic behavior of the tearing parity layer solutions is such that 
\begin{equation}\label{e47}
\tilde{\psi}\rightarrow  \psi_0\!\left[1+ \frac{\hat{\mit\Delta}}{2}\,|X| + {\cal O}(X^2)\right]
\end{equation}
as $|X|\rightarrow\infty$, where $\psi_0$ is an arbitrary constant, then asymptotic matching to the outer solution yields
\begin{equation}\label{e48}
{\mit\Delta\Psi}_s= S^{1/3}\,\hat{\mit\Delta}\,\,{\mit\Psi}_s.
\end{equation}

\section{Linear Resonant Plasma Response Regimes}\label{linear}
\subsection{Introduction}
The aim of this section is to solve the linear layer equations derived in Sect.~\ref{layer} to determine the
response of a resonant layer to an external perturbation.

\subsection{Fourier Transformation}
Equations~(\ref{e43})--(\ref{e46}) are most conveniently solved in Fourier transform space.\cite{cole} 
Let
\begin{equation}
\bar{\phi}(p) = \int_{-\infty}^\infty \tilde{\phi}(X)\,{\rm e}^{-{\rm i}\,p\,X}\,dX,
\end{equation}
et cetera. The Fourier transformed linear layer equations become
\begin{align}\label{e314}
-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_{e}]\,\bar{\psi}&=\frac{d}{dp}\!\left[\bar{\phi}-(1+\lambda_e)\,\bar{N}\right] -p^2\,\bar{\psi},\\[0.5ex]
-{\rm i}\,(Q-Q_E -Q_{i})\,p^2\,\bar{\phi}&=  \frac{d(p^2\,\bar{\psi})}{dp}- P_\varphi\,p^4\!\left(\bar{\phi} + \frac{\bar{N}}{\tau}\right),\\[0.5ex]
-{\rm i}\,(Q-Q_E)\bar{N} &= - {\rm i}\,Q_{e}\,\bar{\phi} + c_\beta^{\,2}\,\frac{d\bar{V}}{dp}-D^{2}\,\frac{d(p^2\,\bar{\psi})}{dp}
  - P_\perp\,p^2\bar{N},\label{e316}\\[0.5ex]
 -{\rm i}\,(Q-Q_E)\,\bar{V} &= {\rm i}\,Q_{e}\,\bar{\psi} +\frac{d\bar{N}}{dp}- P_\varphi\,p^2\,\bar{V},\label{e317}
\end{align}
where, for a tearing parity solution, 
\begin{equation}
\bar{\phi}(p)\rightarrow \bar{\phi}_0\!\left[\frac{\hat{\mit\Delta}}{\pi\,p} + 1+ {\cal O}(p)\right]
\end{equation}
as $p\rightarrow 0$.

Let us ignore the term $c_\beta^{\,2}\,d\bar{V}/dp$ in Eq.~(\ref{e316}). This approximation can
be justified {\em a posteriori}. It is equivalent to the neglect of the contribution of ion parallel dynamics to the linear  plasma
response in the resonant layer, and effectively decouples Eq.~(\ref{e317}) from Eqs.~(\ref{e314})--(\ref{e316}).\cite{cole,ber} Equations~(\ref{e314})--(\ref{e316}) reduce to\,\cite{cole}
\begin{equation}\label{e320}
\frac{d}{dp}\!\left[G(p)\,\frac{dY_e}{dp}\right] - \frac{A(p)}{B(p)}\,p^{2}\,Y_e = 0,
\end{equation}
where
\begin{align}
G(p)&= \frac{p^2}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] + p^2},\\[0.5ex]
 A(p) &= -(Q-Q_E)\,(Q-Q_E-Q_i)-{\rm i}\,(Q-Q_E-Q_i)\,(P_\varphi+P_\perp)\,p^2 + P_\varphi\,P_\perp\,p^4,\\[0.5ex]
 B(p) &= 
-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] +\{P_\perp - {\rm i}\,(1+\lambda_e)\,(Q-Q_E-Q_i)\,D^2\}p^2\nonumber\\[0.5ex]
&\phantom{=}+[(1+\lambda_e)+1/\tau]\,P_\varphi\,D^2\,p^4,
\end{align}
and $Y_e=\bar{\phi} - (1+\lambda_e)\,\bar{N}$. The boundary conditions are that $Y_e(p)$ is bounded as $p\rightarrow\infty$, and
\begin{equation}\label{e325}
Y_e(p)\rightarrow Y_0\!\left[\frac{\hat{\mit\Delta}}{\pi\,p} + 1+ {\cal O}(p)\right]
\end{equation}
as $p\rightarrow 0$. In the following, we shall assume that $|Q|\sim |Q_E|\sim |Q_e|\sim |Q_i|$, $P_\varphi\sim P_\perp \sim P$, $\tau\sim {\cal O}(1)$, and $\lambda_e\sim{\cal O}(1)$,  for the
sake of simplicity.

\subsection{Constant-$\psi$ Approximation}\label{s5.6}
Let us suppose that there are two layers in $p$ space. In the small-$p$ layer, suppose that Eq.~(\ref{e320}) reduces to
\begin{equation}
\frac{d}{dp}\!\left[\frac{p^2}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e] + p^2}\,\frac{dY_e}{dp}\right]\simeq 0
\end{equation}
when $p\sim Q^{1/2}$. Integrating directly, we find that
\begin{equation}
Y_e(p) \simeq Y_0 \left[\frac{\hat{\mit\Delta}}{\pi}\left(\frac{1}{p} + \frac{p}{{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}\right) + 1 + {\cal O}(p^2)\right]
\end{equation}
for $p\lesssim {\cal O}(Q^{1/2})$, where use has been made of Eq.~(\ref{e325}). The two-layer  approximation is
equivalent to the well-known {\em constant-$\psi$ approximation}.\cite{fkr}

In the large-$p$ layer, for  $p\gg {\cal O}(Q^{1/2})$, we obtain 
\begin{equation}\label{e329}
\frac{d^{\,2} Y_e}{dp^2} - \frac{A(p)}{B(p)}\,p^2\,Y_e\simeq 0,
\end{equation}
with $Y_e$ bounded as $p\rightarrow \infty$. Asymptotic matching to the small-$p$ layer yields the boundary
condition 
\begin{equation}\label{e330}
Y_e(p)\simeq Y_0\left\{1+ \frac{\hat{\mit\Delta}}{\pi}\,\frac{p}{{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}+ {\cal O}(p^2)\right\}
\end{equation}
as $p\rightarrow 0$.

In the various constant-$\psi$ linear response regimes considered in Sect.~\ref{scp}, Eq.~(\ref{e329}) reduces to an
equation of the form
\begin{equation}\label{e330a}
\frac{d^{\,2}Y_e}{dp^2}- C\,p^m\,Y_e\simeq 0,
\end{equation}
where $m$ is real and non-negative, and $C$ is a complex constant. Let $Y_e=\sqrt{p}\,Z$ and $q=\sqrt{C}\,p^n/n$, where
$n=(m+2)/2$. The previous equation transforms into a modified Bessel equation of general order, 
\begin{equation}
q^2\,\frac{d^{\,2}Z}{dq^2} + q\,\frac{dZ}{dq} - (q^2+\nu^2)\,Z =0,
\end{equation}
where $\nu=1/(m+2)$. The solution that is bounded as $q\rightarrow \infty$ has the small-$q$
expansion
\begin{equation}\label{e332}
K_\nu(z)= \frac{1}{{\mit\Gamma}(1-\nu)}\left(\frac{2}{q}\right)^\nu - \frac{1}{{\mit\Gamma}(1+\nu)}\left(\frac{q}{2}\right)^\nu + {\cal O}(q^{2-\nu}),
\end{equation}
where ${\mit\Gamma}(z)$ is a gamma function. 
A comparison of this expression with Eq.~(\ref{e330}) reveals that
\begin{equation}
\hat{\mit\Delta} = \pi\,(-{\rm i}\,\left[Q-Q_E-(1+\lambda_e)\,Q_e\right])\, \frac{{\mit\Gamma}(1-\nu)}{{\mit\Gamma}(\nu)}\,\nu^{2\nu-1}\,C^{\,\nu}.
\end{equation}
Note, finally, that $p_\ast\sim |C|^{-\nu}$, where $p_\ast$ denotes the width of the large-$p$ layer in $p$ space. This width must be larger than $Q^{1/2}$ (i.e., the width of the small-$p$ layer) in order for the constant-$\psi$ approximation to hold.

\subsection{Constant-$\psi$ Linear Response Regimes}\label{scp}
Suppose that $Q\gg P\,p^2$ and $D^2\,p^2 \ll 1$. It follows that $\nu=1/4$ and
\begin{equation}
C = \frac{[-{\rm i}\,(Q-Q_E)]\,[-{\rm i}\,(Q-Q_E-Q_i)]}{-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e]}.
\end{equation}
Hence, we deduce that
\begin{equation}\label{ri}
\hat{\mit\Delta} = \frac{2\pi\,{\mit\Gamma}(3/4)}{{\mit\Gamma}(1/4)}\,(-{\rm i}\,[Q-Q_E-(1+\lambda_e)\,Q_e])^{3/4}\,[-{\rm i}\,(Q-Q_E)]^{1/4}\,[-{\rm i}\,(Q-Q_E-Q_i)]^{1/4}.
\end{equation}
This response regime is known as the {\em resistive-inertial regime}, because the layer response is dominated by
plasma resistivity and ion inertia.\cite{fkr,ara,rf1998}. The characteristic layer width is $p_\ast \sim Q^{-1/4}$,
which implies that the regime is valid when $P\ll Q^{3/2}$, $Q\gg D^4$, and $Q\ll 1$.
